Convergent Evolution During Local Adaptation to Patchy Landscapes

### 2014-07-11 patchy-selection-paper.tex
plos reference style

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{color}
\usepackage{natbib}
\usepackage[hidelinks]{hyperref}

@@

% \usepackage{cite}

% Use doublespacing - comment out for single spacing
%\usepackage{setspace} 
%\doublespacing

@@

% turn this on to doublespace and put all figures at the very end.
\newif\ifsubmissionversion
\submissionversionfalse
\newif\ifincludesupplement
\includesupplementfalse

\ifsubmissionversion
\usepackage{endfloat}
\fi

\ifsubmissionversion
% \linespread{1.6}
\doublespacing
\fi

% Use the PLoS provided bibtex style
% OR NOT \bibliographystyle{plos2009}
\bibliographystyle{abbrvnat}

@@

\paragraph{Past work on spatial models}
We will make use of several previous results from the literature on
migration in models of spatially varying selection.
\citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
deterministic models of of this sort.
\citet{nagylaki1975conditions} and \citet{conley1975application} showed that in one dimension, if the physical width of the patch is less than $(\sigma/\sqrt{2s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
then there is no stable equilibrium with both $b$ and $B$ present,  % see p.603 in Nagylaki
so that migrational swamping prevents $B$ from establishing \citep[see also][ for a review]{lenormand2002limits}.
\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed  
that patches must be at least this critical size so that a new mutation has positive probability of establishment
(in the large population density limit).
The same paper also showed that
this probability of establishment of a new mutation decays exponentially with distance from the patch, 
with the scale given by $\sigma/\sqrt{2s_m}$. 
Mutations appearing within the patch have probability of establishment
strictly less than the probability for a panmictic population,
but approaching this as the size of the patch increases.
This latter panmictic probability of establishment we denote $p_e$,
and often approximate $p_e$ by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
This result holds quite generally for a geographically spread population that experiences a uniform selection
pressure \citep{maruyama1970fixation,cherry2003diffusion}. 
More general work on the mathematically equivalent problem of density-dependent population growth 
has obtained much more general criteria
under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1989diffusive},
and determined the habitat configuration that maximizes the probability of establishment \citep{lou2006minimization}.

@@

Since each individual has on average $e^{-s_m}$ offspring,
the expected family size after time $t$ is $e^{- t s_m}$.
Therefore, if we define $k_e(t)$ to be the chance that the family
has gone extinct by time $t$, 
and let $K_t$ be the (random) family size conditioned on nonextinction,
then by Bayes' rule, $e^{-s_m t} = (1-k_e(t))\E[K_t]$.
It turns out that if the distribution of offspring numbers is not too heavy-tailed \citep[see][for details]{jagers1975branching},
then $K_t$ has a limiting distribution: $K_t \dconv K$,
and $e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
In other words, the chance of survival to time $t$ is a constant multiplied by $e^{-s_m t}$,
and conditional on survival, the family size is given by a fixed distribution.

This suggests that families can be well-described by motion of a central ``trunk'',
which dies at a random time whose distribution is given by $k_e(t)$,
and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
as depicted in Figure~\ref{fig:branching_decomp}.
We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} \citep[see also][]{chauvin1991growing}.
When we condition on survival until $t$, we condition on existence of
at least one trunk lineage surviving from time $0$ to time $t$ (the long red
lineage in Figure \ref{fig:branching_decomp}).
Given this lineage, nothing distinguishes any of the other reproduction events --
each individual along the trunk lineage must give birth to at least one offspring,
and all other reproduction events are unconditioned.
In other words, the genealogy of a family that survives until $t$
can be decomposed into a trunk that lasts until $t$
and that sprouts independent branches that may or may not survive until $t$.
Specifically, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
% (so $\P\{Z_t = 0\} = k_e(t)$),
then this gives the following decomposition:
\begin{align}
  \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
\end{align}
where each $Z^{(s,k)}$ is an independent copy of $Z$ 
that arose from the main branch in generation $s$,
and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.

@@

\begin{figure}[ht]
  \begin{center}
    \includegraphics{sim-transit.pdf}
  \end{center}
  \caption{
  \textbf{Adaptation by migration:}
  An initially unadapted patch (10 demes at the left end of the range) is colonized around generation 5000;
  a representative lineage is shown in the new patch, which follows the original migrant ``trunk'' lineage
  between the patches and coalesces with a lineage in the unadapted patch.
  In the left panel, as in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
  other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
  The right panel shows (in red) the extent of the shared haplotype between the two lineages around the selected locus (at position 0),
  which decreases as time progresses. 
  \label{fig:lineagesmotion}
  }
\end{figure}

@@

If a patch adapts through new mutation or a rare migrant lineage, the
genomic region surrounding the selected site will hitchhike along with it \citep{maynardsmith1974hitchhiking},
so that at least initially, all adapted individuals within a patch
share a fairly long stretch of haplotype. 
Pairs of adapted individuals within a patch will share a haplotype of mean genetic length of
about $\log(\rho A s_p)/s_p$ around the selected site, 
as long as the patch is reasonably well mixed by dispersal
\citep[otherwise see][]{barton2013genetic}.
This association gets slowly whittled down by recombination during interbreeding at the edge of the patch,
but there will always be longer LD nearby to the selected site.

@@

Our first obstacle is that we do not know $1-k_e(t)$, 
the probability a family survives until $t$, exactly.
However, we do know that $1-k_e(t) \simeq e^{-s_m t} / \E[K]$,
i.e.\ that it has exponential tails.
Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
supposing that the random lifetime of a family is exponentially distributed with mean $1/s_m$.
Once we do this, 
we can use standard results on hitting times of $d$-dimensional Brownian motion
that is killed at rate $s_m$ \citep[see][2.2.0.1 and 4.2.0.1]{borodin2002handbook}.
In particular, if the patch is circular with radius $w$ and lies at distance
$R$ from the already adapted patch, then 
we know that the Laplace transform of the transit time $\tau$ is
\begin{align}
  \E[e^{-\ell \tau}] &=
    \begin{cases}
      e^{- R \sqrt{2(s_m+\ell)}/\sigma} \qquad & d=1 \\
      \frac{ K_0( (R+w)\sqrt{2(s_m+\ell)}/\sigma) }{ K_0( w\sqrt{2(s_m+\ell)}/\sigma) } \qquad & d=2  ,
    \end{cases} \label{eqn:borodinresult}
    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(s_m+\ell)}/\sigma}  (1+O(1/w)) .
\end{align}
where $K_0$ is a modified Bessel function of the second kind.
We are interested in lineages that manage to reach the patch before being killed,
i.e.\ having $\tau < \infty$,
which occurs with probability
$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau) \right]$.

@@

A detailed treatment of this is well beyond the scope of this paper,
but we can get a good idea of the genomic scale of these effects through heuristic arguments.
To understand the probability of a recombination moving our lineage
off of the B background we need to understand what fraction of the
time our lineage spends in geographic regions where $b$ is common. 
We will follow the lineage of a single $B$ allele
in a patch at migration-selection balance, 
i.e.\ with the local proportion of $B$ alleles near $x$ equal to $q(x)$,
and suppose that the lineage stays close enough to the patch 
that fluctuations about $q(x)$ are small. 
(Note that this assumption differs from
previous sections, but here our focus is on a typical lineage 
in the patch rather than the rare lineages that make it far from the patch).
If we suppose that type $B$ individuals at location $y$ reproduce at rate $1+s(y)$,
and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
then the total rate of influx of $B$ alleles from $y$ into $x$ is $\rho q(y) (1+s(y)) m(y,x)$.
As argued by \citet{hudson1988coalescent}, 
this implies that the chance that a $B$ allele at $x$
has been inherited from a parent at $y$ in a time interval $dt$
is equal to this influx divided by the local abundance of $B$ alleles,
i.e.\ $dt \times q(y) (1+s(y)) m(y,x) / q(x)$.
This therefore describes the movement of a lineage backwards in time.
After the same large-population, continuous-space limit that underlies the rest of this paper (section~\nameref{apx:eqfreq}),
if the assumption that fluctuations in $q(x)$ are small,
this motion should converge to a diffusion process \citep[as assumed by e.g.][]{hallatschek2008surfing},
notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
This diffusion, when at $x$, 
is pulled toward regions with more $B$ alleles
by a mean displacement term $(1+s(x))\frac{d}{d\,x} \log( q(x) (1+s(x)) )$,
where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
and moves at speed $\sqrt{1+s(x)}$ (which is nearly constant, since $s(x)$ is small).
In It\=o differential notation, 
the location $X_t$ of a $B$ lineage $t$ generations ago satisfies
\begin{align}
    d X_t &= (1+s(X_t)) \frac{d}{dx} \log( q(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,
\end{align}
where $B_t$ is a standard Brownian motion.
Furthermore, it is easy to verify that the stationary distribution of this process has density
\begin{align}
  \pi(x) = \frac{ q(x)^2 (1+s(x)) }{ \int_y q(y)^2 (1+s(y)) dy } . \label{eqn:lineagestatdist}
\end{align}
A lineage at a locus at genetic distance $R$ from a $B$ allele
moves as a $B$ lineage until it recombines onto a $b$ background,
which occurs when at $x$ with rate $r (1-q(x)) (1+s(x)/2)$.
The Feynman-Kac formula could then be used to write down differential equations
solved by various moments of the haplotype length as a function of the sampling location of the lineage;
instead, we only use the above as a guide for rough approximation.

@@

A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. 
\citet{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
On the lava flow the dark MC1R allele is at 86\% frequency,
while at a site with light substrate 12km to the west (Tule), the frequency is 3\%.
The dark allele was entirely absent from Christmas pass, a site with light substrate 7km north of Tule, and 3km further from the lava flow.
In the other direction, the dark MC1R allele was at 34\% at a site with light substrate 10km to the east of the flow (O'Neill).
These numbers give us a sense of a plausible range of characteristic lengths.
Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
we can fit formula \eqref{eqn:eqfreq} to these frequencies
\citep[similar to][]{hoekstra2004ecological}.
Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
and for O'Neill, $\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. 
We also need estimate of $s_m$. 
Using $\sigma \approx 1$km \citep{french1968dispersal,allred1963range}, 
these cline widths imply that $s_m=1/9$ and $1/900$ are reasonable values.

@@

This paper is an investigation into the basic question: 
What is the spatial resolution of convergent local adaptation?
In other words, 
over what spatial scale of environmental patchiness will the process
of adaptation develop independent solutions to evolutionary problems?
The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
the dispersal distance divided by the square root of the strength of selection against the allele between patches. 
It depends much more weakly on the
selective benefit within the patches or (perhaps surprisingly) the population density, 
although these two factors will determine the time-scale over which adaptation will occur. 
This is in contrast with models of panmictic populations
\citep{softsweepsI, MesserPetrov, Wilson-softsweep} 
and geographically spread populations adapting to homogeneous selection pressures \citep{ralph2010parallel}, 
where the probability of multiple, independently arising adaptive alleles increases with the population size/density. 
However, in all of these models the dependence on the beneficial selection
coefficient is absent or weak, due to the fact that selection both
aids establishment of new alleles and the spread of existing
alleles \citep[but see][for the complications of varying population sizes]{Wilson-softsweep}.

@@

\paragraph{Long distance migration}
We have also ignored the possibility of very long distance migration,
instead focusing on local dispersal (and using a Gaussian model).
However, dispersal distributions can be very heavy tailed, 
with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
In addition, over long time-scales, very rare chance events \citep[mice carried off by hurricanes and the like][]{censky1998overwater,nathan2008mechanisms}
could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
Such tail events could greatly increase the probability of parallel adaptation above that predicted by our model. 
Furthermore, if adaptive alleles do move between distant patches via rare, long distance migration 
then they will be associated with a much longer shared haplotype than predicted by equation \eqref{eqn:haplotype_length}. 
As such, we view our results as a null model by which the contribution of long distribution migrants to adaptation could be empirically judged.

@@

If $q(r)$ is a radially symmetric equilibrium frequency and $\xi(t,x)=q(|x|)$,
then $\partial_t \xi(t,x) = 0$.
So, using \eqref{eqn:radial_fkpp}, a radially symmetric solution 
with $s(r) = -s < 0$ for all $r>r_0$ 
solves, for $r_0 < r < \infty$,
\begin{align} \label{eqn:radial_eqfreq_diffeq}
    \partial_r^2 q(r) + \frac{d-1}{r} \partial_r q(r) - \frac{2s}{\sigma^2} q(r) (1-q(r)) = 0 \\
    \lim_{r \to \infty} q(r) = \lim_{r \to \infty} \partial_r q(r) = 0 \qquad 
    0 < q(r) < 1 
\end{align}
Since $q(r) \to 0$ as $r \to \infty$, so $q(r) (1-q(r)) \approx q(r)$,
it can be shown 
% theory in \citet{wasow1965asymptotic} implies 
that the true equilibrium frequency $q$ is close, for large $r$, to the solution to
\begin{align}
    \partial_r^2 u(r) + \frac{d-1}{r} \partial_r u(r) - \frac{2s}{\sigma^2} u(r) = 0  \label{eqn:bessel} \\
    \lim_{r \to \infty} u(r) = \lim_{r \to \infty} \partial_r u(r) = 0 \qquad  .
\end{align}
This has general solution given by a modified Bessel function:
using \citet{gradshteyn2007table} 8.494.9,
the general solution is
\begin{align}
    u(r) = C' (r-r_0)^{(2-d)/2} K_{(2-d)/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
\end{align}
where $C'$ and $r_0$ are chosen to match boundary conditions.
Asymptotics of Bessel functions \citep[][8.451.6]{gradshteyn2007table} then imply that
\begin{align}
    q(r) \approx C r^{(1-d)/2} \exp \left( -r \sqrt{2s}/\sigma \right) + O(1/r),
\end{align}
where $C$ is a different constant.

>>>>

\documentclass{article}
\usepackage{fullpage}
\usepackage{amsmath,amssymb}
\usepackage{marginnote}
\usepackage{graphicx}
\usepackage{color}
\usepackage[hidelinks]{hyperref}

@@

@@

% turn this on to doublespace and put all figures at the very end.
\newif\ifsubmissionversion
% \submissionversiontrue
\submissionversionfalse

\ifsubmissionversion
    \usepackage{endfloat}
    \usepackage{cite}
    % Use the PLoS provided bibtex style
    \bibliographystyle{plos2009}
    \newcommand{\citep}[1]{\cite{#1}}
    \newcommand{\citet}[1]{\cite{#1}}
\else
    \usepackage{natbib}
    \bibliographystyle{abbrvnat}
\fi

\ifsubmissionversion
% \linespread{1.6}
%\usepackage{setspace} 
%\doublespacing
\fi

@@

\paragraph{Past work on spatial models}
We will make use of several previous results from the literature on
migration in models of spatially varying selection.
\citet{haldane1948theory}, \citet{fisher1950frequencies}, and later \citet{slatkin1973geneflow} first analyzed
deterministic models of of this sort.
\citet{nagylaki1975conditions} and \citet{conley1975application} showed that in one dimension, if the physical width of the patch is less than $(\sigma/\sqrt{2s_p}) \tan^{-1} (\sqrt{s_m/s_p})$, 
then there is no stable equilibrium with both $b$ and $B$ present,  % see p.603 in Nagylaki
so that migrational swamping prevents $B$ from establishing (see also \citet{lenormand2002limits} for a review).
\citet{barton1987establishment}, using general theory of \citet{pollak1966survival}, showed  
that patches must be at least this critical size so that a new mutation has positive probability of establishment
(in the large population density limit).
The same paper also showed that
this probability of establishment of a new mutation decays exponentially with distance from the patch, 
with the scale given by $\sigma/\sqrt{2s_m}$. 
Mutations appearing within the patch have probability of establishment
strictly less than the probability for a panmictic population,
but approaching this as the size of the patch increases.
This latter panmictic probability of establishment we denote $p_e$,
and often approximate $p_e$ by $2 s_p / \xi^2$ \citep{haldane1927mathematical,fisher1930genetical}.
This result holds quite generally for a geographically spread population that experiences a uniform selection
pressure \citep{maruyama1970fixation,cherry2003diffusion}. 
More general work on the mathematically equivalent problem of density-dependent population growth 
has obtained much more general criteria
under which a habitat configuration will support a stable polymorphic equilibrium \citep{cantrell1989diffusive},
and determined the habitat configuration that maximizes the probability of establishment \citep{lou2006minimization}.

@@

Since each individual has on average $e^{-s_m}$ offspring,
the expected family size after time $t$ is $e^{- t s_m}$.
Therefore, if we define $k_e(t)$ to be the chance that the family
has gone extinct by time $t$, 
and let $K_t$ be the (random) family size conditioned on nonextinction,
then by Bayes' rule, $e^{-s_m t} = (1-k_e(t))\E[K_t]$.
It turns out that if the distribution of offspring numbers is not too heavy-tailed (see \citet{jagers1975branching} for details),
then $K_t$ has a limiting distribution: $K_t \dconv K$,
and $e^{-s_m t}/(1-k_e(t)) \to \E[K]$ as $t \to \infty$.
In other words, the chance of survival to time $t$ is a constant multiplied by $e^{-s_m t}$,
and conditional on survival, the family size is given by a fixed distribution.

This suggests that families can be well-described by motion of a central ``trunk'',
which dies at a random time whose distribution is given by $k_e(t)$,
and conditional on survival of this trunk, the family is composed of side ``branches'' off of this trunk,
as depicted in Figure~\ref{fig:branching_decomp}.
We can understand this quite concretely, by a method described rigorously by \citet{geiger1999elementary} (see also \citet{chauvin1991growing}).
When we condition on survival until $t$, we condition on existence of
at least one trunk lineage surviving from time $0$ to time $t$ (the long red
lineage in Figure \ref{fig:branching_decomp}).
Given this lineage, nothing distinguishes any of the other reproduction events --
each individual along the trunk lineage must give birth to at least one offspring,
and all other reproduction events are unconditioned.
In other words, the genealogy of a family that survives until $t$
can be decomposed into a trunk that lasts until $t$
and that sprouts independent branches that may or may not survive until $t$.
Specifically, if we write $Z_t$ for the number of offspring in the branching process alive at time $t$,
% (so $\P\{Z_t = 0\} = k_e(t)$),
then this gives the following decomposition:
\begin{align}
  \left( Z_t \; \vert \; Z_t>0 \right) \deq 1 + \sum_{s=0}^t \sum_{k=1}^{W_s-1} Z^{(s,k)}_{t-s},
\end{align}
where each $Z^{(s,k)}$ is an independent copy of $Z$ 
that arose from the main branch in generation $s$,
and each $W_s$ is an independent copy of the offspring distribution, conditioned on $W_s \ge 1$.

@@

\begin{figure}[ht]
  \begin{center}
      \includegraphics{sim-transit}
  \end{center}
  \caption{
  \textbf{Adaptation by migration:}
  An initially unadapted patch (10 demes at the left end of the range) is colonized around generation 5000;
  a representative lineage is shown in the new patch, which follows the original migrant ``trunk'' lineage
  between the patches and coalesces with a lineage in the unadapted patch.
  In the left panel, as in Figure~\ref{fig:sim_snapshots}, color and contours show allele frequency across space and time;
  other simulation parameters are as in Figures~\ref{fig:sim_occupation_freqs}B and \ref{fig:sim_snapshots}.
  The right panel shows (in red) the extent of the shared haplotype between the two lineages around the selected locus (at position 0),
  which decreases as time progresses. 
  \label{fig:lineagesmotion}
  }
\end{figure}

@@

If a patch adapts through new mutation or a rare migrant lineage, the
genomic region surrounding the selected site will hitchhike along with it \citep{maynardsmith1974hitchhiking},
so that at least initially, all adapted individuals within a patch
share a fairly long stretch of haplotype. 
Pairs of adapted individuals within a patch will share a haplotype of mean genetic length of
about $\log(\rho A s_p)/s_p$ around the selected site, 
as long as the patch is reasonably well mixed by dispersal
(otherwise see \citet{barton2013genetic}).
This association gets slowly whittled down by recombination during interbreeding at the edge of the patch,
but there will always be longer LD nearby to the selected site.

@@

Our first obstacle is that we do not know $1-k_e(t)$, 
the probability a family survives until $t$, exactly.
However, we do know that $1-k_e(t) \simeq e^{-s_m t} / \E[K]$,
i.e.\ that it has exponential tails.
Since $\E[K] \approx 1/s_m$, it is sensible to make one further approximation,
supposing that the random lifetime of a family is exponentially distributed with mean $1/s_m$.
Once we do this, 
we can use standard results on hitting times of $d$-dimensional Brownian motion
that is killed at rate $s_m$ (see \citet{borodin2002handbook} 2.2.0.1 and 4.2.0.1).
In particular, if the patch is circular with radius $w$ and lies at distance
$R$ from the already adapted patch, then 
we know that the Laplace transform of the transit time $\tau$ is
\begin{align}
  \E[e^{-\ell \tau}] &=
    \begin{cases}
      e^{- R \sqrt{2(s_m+\ell)}/\sigma} \qquad & d=1 \\
      \frac{ K_0( (R+w)\sqrt{2(s_m+\ell)}/\sigma) }{ K_0( w\sqrt{2(s_m+\ell)}/\sigma) } \qquad & d=2  ,
    \end{cases} \label{eqn:borodinresult}
    % &= \left( R/w +1 \right)^{-\frac{d-1}{2}} e^{- R \sqrt{2(s_m+\ell)}/\sigma}  (1+O(1/w)) .
\end{align}
where $K_0$ is a modified Bessel function of the second kind.
We are interested in lineages that manage to reach the patch before being killed,
i.e.\ having $\tau < \infty$,
which occurs with probability
$\P\{\tau < \infty\} = \lim_{\ell \to 0} \E \left[\exp(-\ell \tau) \right]$.

@@

A detailed treatment of this is well beyond the scope of this paper,
but we can get a good idea of the genomic scale of these effects through heuristic arguments.
To understand the probability of a recombination moving our lineage
off of the B background we need to understand what fraction of the
time our lineage spends in geographic regions where $b$ is common. 
We will follow the lineage of a single $B$ allele
in a patch at migration-selection balance, 
i.e.\ with the local proportion of $B$ alleles near $x$ equal to $q(x)$,
and suppose that the lineage stays close enough to the patch 
that fluctuations about $q(x)$ are small. 
(Note that this assumption differs from
previous sections, but here our focus is on a typical lineage 
in the patch rather than the rare lineages that make it far from the patch).
If we suppose that type $B$ individuals at location $y$ reproduce at rate $1+s(y)$,
and the chance their offspring move to $x$ is $m(y,x)$ (with $\sum_x m(y,x)=1$),
then the total rate of influx of $B$ alleles from $y$ into $x$ is $\rho q(y) (1+s(y)) m(y,x)$.
As argued by \citet{hudson1988coalescent}, 
this implies that the chance that a $B$ allele at $x$
has been inherited from a parent at $y$ in a time interval $dt$
is equal to this influx divided by the local abundance of $B$ alleles,
i.e.\ $dt \times q(y) (1+s(y)) m(y,x) / q(x)$.
This therefore describes the movement of a lineage backwards in time.
After the same large-population, continuous-space limit that underlies the rest of this paper (section~\nameref{apx:eqfreq}),
if the assumption that fluctuations in $q(x)$ are small,
this motion should converge to a diffusion process (as assumed by e.g.\ \citet{hallatschek2008surfing}),
notwithstanding some technical difficulties (e.g.\ discontinuity of $s(x)$).
This diffusion, when at $x$, 
is pulled toward regions with more $B$ alleles
by a mean displacement term $(1+s(x))\frac{d}{d\,x} \log( q(x) (1+s(x)) )$,
where $s(x)=s_p$ if $x$ is in the patch and $s(x)=s_m$ otherwise, 
and moves at speed $\sqrt{1+s(x)}$ (which is nearly constant, since $s(x)$ is small).
In It\=o differential notation, 
the location $X_t$ of a $B$ lineage $t$ generations ago satisfies
\begin{align}
    d X_t &= (1+s(X_t)) \frac{d}{dx} \log( q(X_t) (1+s(X_t)) ) dt + \sigma \sqrt{1+s(X_t)} dB_t,
\end{align}
where $B_t$ is a standard Brownian motion.
Furthermore, it is easy to verify that the stationary distribution of this process has density
\begin{align}
  \pi(x) = \frac{ q(x)^2 (1+s(x)) }{ \int_y q(y)^2 (1+s(y)) dy } . \label{eqn:lineagestatdist}
\end{align}
A lineage at a locus at genetic distance $R$ from a $B$ allele
moves as a $B$ lineage until it recombines onto a $b$ background,
which occurs when at $x$ with rate $r (1-q(x)) (1+s(x)/2)$.
The Feynman-Kac formula could then be used to write down differential equations
solved by various moments of the haplotype length as a function of the sampling location of the lineage;
instead, we only use the above as a guide for rough approximation.

@@

A key parameter above was the dispersal distance divided by the square root of strength of selection against the focal allele between patches, $\sigma/\sqrt{s_m}$. 
\citet{hoekstra2004ecological} studied the frequency of the dark MC1R allele and coat color phenotypes, 
at sites across the (dark) Pinacate lava flow and at two nearby sites in light-colored rock.
On the lava flow the dark MC1R allele is at 86\% frequency,
while at a site with light substrate 12km to the west (Tule), the frequency is 3\%.
The dark allele was entirely absent from Christmas pass, a site with light substrate 7km north of Tule, and 3km further from the lava flow.
In the other direction, the dark MC1R allele was at 34\% at a site with light substrate 10km to the east of the flow (O'Neill).
These numbers give us a sense of a plausible range of characteristic lengths.
Assuming the dark allele is at 50\% frequency at the edge of the lava flow, 
we can fit formula \eqref{eqn:eqfreq} to these frequencies
(similar to \citet{hoekstra2004ecological}).
Doing this for Tule we obtain $\sigma/\sqrt{s_m} \approx 3$km, 
and for O'Neill, $\sigma/\sqrt{s_m} \approx 30$km, giving us a range of cline widths. 
We also need estimate of $s_m$. 
Using $\sigma \approx 1$km \citep{french1968dispersal,allred1963range}, 
these cline widths imply that $s_m=1/9$ and $1/900$ are reasonable values.

@@

This paper is an investigation into the basic question: 
What is the spatial resolution of convergent local adaptation?
In other words, 
over what spatial scale of environmental patchiness will the process
of adaptation develop independent solutions to evolutionary problems?
The answer to this depends most strongly on $\sigma/\sqrt{s_m}$, 
the dispersal distance divided by the square root of the strength of selection against the allele between patches. 
It depends much more weakly on the
selective benefit within the patches or (perhaps surprisingly) the population density, 
although these two factors will determine the time-scale over which adaptation will occur. 
This is in contrast with models of panmictic populations
\citep{softsweepsI, MesserPetrov, Wilson-softsweep} 
and geographically spread populations adapting to homogeneous selection pressures \citep{ralph2010parallel}, 
where the probability of multiple, independently arising adaptive alleles increases with the population size/density. 
However, in all of these models the dependence on the beneficial selection
coefficient is absent or weak, due to the fact that selection both
aids establishment of new alleles and the spread of existing
alleles (but see \citet{Wilson-softsweep} for the complications of varying population sizes).

@@

\paragraph{Long distance migration}
We have also ignored the possibility of very long distance migration,
instead focusing on local dispersal (and using a Gaussian model).
However, dispersal distributions can be very heavy tailed, 
with a small fraction of individuals moving very long distances indeed \citep{levin2003ecology,reynolds2009levy}.
In addition, over long time-scales, very rare chance events (mice carried off by hurricanes and the like \citet{censky1998overwater,nathan2008mechanisms})
could play a role in spreading migrant alleles if adaptation by other means is sufficiently unlikely.
Such tail events could greatly increase the probability of parallel adaptation above that predicted by our model. 
Furthermore, if adaptive alleles do move between distant patches via rare, long distance migration 
then they will be associated with a much longer shared haplotype than predicted by equation \eqref{eqn:haplotype_length}. 
As such, we view our results as a null model by which the contribution of long distribution migrants to adaptation could be empirically judged.

@@

If $q(r)$ is a radially symmetric equilibrium frequency and $\xi(t,x)=q(|x|)$,
then $\partial_t \xi(t,x) = 0$.
So, using \eqref{eqn:radial_fkpp}, a radially symmetric solution 
with $s(r) = -s < 0$ for all $r>r_0$ 
solves, for $r_0 < r < \infty$,
\begin{align} \label{eqn:radial_eqfreq_diffeq}
    \partial_r^2 q(r) + \frac{d-1}{r} \partial_r q(r) - \frac{2s}{\sigma^2} q(r) (1-q(r)) = 0 \\
    \lim_{r \to \infty} q(r) = \lim_{r \to \infty} \partial_r q(r) = 0 \qquad 
    0 < q(r) < 1 
\end{align}
Since $q(r) \to 0$ as $r \to \infty$, so $q(r) (1-q(r)) \approx q(r)$,
it can be shown 
% theory in \citet{wasow1965asymptotic} implies 
that the true equilibrium frequency $q$ is close, for large $r$, to the solution to
\begin{align}
    \partial_r^2 u(r) + \frac{d-1}{r} \partial_r u(r) - \frac{2s}{\sigma^2} u(r) = 0  \label{eqn:bessel} \\
    \lim_{r \to \infty} u(r) = \lim_{r \to \infty} \partial_r u(r) = 0 \qquad  .
\end{align}
This has general solution given by a modified Bessel function:
using \citet{gradshteyn2007table} 8.494.9,
the general solution is
\begin{align}
    u(r) = C' (r-r_0)^{(2-d)/2} K_{(2-d)/2} \left( (r-r_0) \sqrt{2s}/\sigma \right) ,
\end{align}
where $C'$ and $r_0$ are chosen to match boundary conditions.
Asymptotics of Bessel functions (\citet{gradshteyn2007table}, 8.451.6) then imply that
\begin{align}
    q(r) \approx C r^{(1-d)/2} \exp \left( -r \sqrt{2s}/\sigma \right) + O(1/r),
\end{align}
where $C$ is a different constant.
